Outliers and discrepancies

### 2024-11-20 misc/indiv.qmd
add manuscript, and restructure website

---
title: "Response Patterns"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  # hugo-md:
  #   include: true
  #   html-math-method: mathjax
  #   output-file: results_hugo.md
  # gfm:
  #   echo: true
  #   output-file: results_gfm.md
---

>>>>

---
title: "Response Patterns"
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  # hugo-md:
  #   include: true
  #   html-math-method: mathjax
  #   output-file: results_hugo.md
  # gfm:
  #   echo: true
  #   output-file: results_gfm.md
---

### 2024-11-20 misc/ordinal.qmd
add manuscript, and restructure website

---
title: "Ordinal Models"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
#lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: ordinal_hugo.md
  gfm:
    echo: true
    output-file: ordinal.md
---

@@

pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())
options(brms.backend="cmdstanr",mc.cores=4)

s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))

@@

library(tidybayes)

# Get predicted probabilities
pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, .category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class")

ggplot(pred_summary, 
        aes(x = refClass, y = mean_prob, fill = .category )) +
    geom_col(position = position_dodge(width = 0.8)) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
                    position = position_dodge(width = 0.8)) +
    labs(x = "Reference Class", 
        y = "Predicted Probability",
        fill = "Accuracy Level",
        title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
    scale_fill_brewer(palette = "RdYlBu") +
    theme_minimal()

# Calculate cumulative probabilities
cum_probs <- pred_summary |>
  group_by(refClass) |>
  arrange(refClass, .category) |>
  mutate(
    cum_prob = cumsum(mean_prob),
    cum_lower = cumsum(lower_ci),
    cum_upper = cumsum(upper_ci)
  )

ggplot(cum_probs, 
    aes(x = refClass, y = cum_prob, color = .category)) +
geom_point(size = 3,position=position_dodge()) +
geom_errorbar(aes(ymin = cum_lower, ymax = cum_upper), 
            width = 0.2,position=position_dodge()) +
labs(x = "Reference Class",
    y = "Cumulative Probability",
    color = "Accuracy Level",
    title = "Cumulative Probabilities of Accuracy Levels") +
theme_minimal() +
scale_color_brewer(palette = "Blues")

summary_table <- pred_summary |>
  pivot_wider(
    names_from = .category,
    values_from = c(mean_prob, lower_ci, upper_ci)
  ) |>
  kable(digits = 3)

print(summary_table)

@@

@@

@@

bayesplot::color_scheme_set(wes_palettes[1])

@@

f |> str()

f <-
  rbind(
    f[, , 1],
    f[, , 2],
    f[, , 3]
  ) |> data.frame() |> bind_cols(expand_grid(new_data, Y=1:3))

@@

@@

library(tidybayes)
library(stringr)

@@

# Calculate Odds Ratios
odds_ratios_s1 <- posterior_samples(ordinal_model_s1) %>%
  mutate(
    OR_Percentage = exp(`b_refClassPercentage`),
    OR_USD = exp(`b_refClassUSD`)
  ) %>%
  select(OR_Percentage, OR_USD) %>%
  summarise_all(list(mean = mean, lower = ~ quantile(., 0.025), upper = ~ quantile(., 0.975))) %>%
  pivot_longer(cols = everything(),
               names_to = c("Reference Class", ".value"),
               names_pattern = "OR_(.*)_(.*)") %>%
  mutate(
    Reference_Class = case_when(
      str_detect(name, "Percentage") ~ "Percentage",
      str_detect(name, "USD") ~ "USD",
      TRUE ~ "kWh"
    )
  )

odds_ratios_s1 <- posterior_summary(ordinal_model_s1, prob = 0.95) %>%
  as.data.frame() %>%
  rownames_to_column(var = "Term") %>%
  filter(str_detect(Term, "refClass")) %>%
  mutate(
    OR = exp(Estimate),
    lower = exp(`Q2.5`),
    upper = exp(`Q97.5`)
  ) %>%
  select(Term, OR, lower, upper)

# Display Odds Ratios
odds_ratios_s1 %>%
  rename(`Reference Class` = Term,
         `Odds Ratio` = OR,
         `95% CrI Lower` = lower,
         `95% CrI Upper` = upper) %>%
  mutate(`Reference Class` = str_replace(`Reference Class`, "refClass", "")) %>%
  kable(booktabs = TRUE, caption = "Study 1: Odds Ratios for Matching Goals by Reference Class") %>%
  kable_styling(full_width = FALSE)

library(caret)

# Create confusion matrix
observed <- s1_agg$accuracy_level
predicted <- predict(ordinal_model_s1, newdata = s1_agg, type = "response")
predicted_class <- apply(predicted, 1, function(x) which.max(x))

confusion_matrix <- confusionMatrix(
  factor(predicted_class, levels = 1:4, labels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error")),
  factor(observed, levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"))
)

# Display Confusion Matrix
confusion_matrix$table %>%
  as.data.frame.matrix() %>%
  rownames_to_column(var = "Predicted") %>%
  kable(booktabs = TRUE, caption = "Study 1: Confusion Matrix for Matched Goals") %>%
  kable_styling(full_width = FALSE)

@@

@@

# compare
loo_compare(loo(ordinal_model_s1), loo(ordinal_model_s1_prb))
compare_models(ordinal_model_s1, ordinal_model_s1_prb)

@@

@@

odds_ratio_summary_s2
#             comparison odds_ratio ci_lower ci_upper
# 1    Percentage vs kWh       1.99     0.73     4.43
# 2           USD vs kWh       3.69     1.33     8.29
# 3       Rounded vs Not       0.70     0.57     0.84
# 4 15% Goal vs 10% Goal       0.78     0.63     0.94

@@

# Fit the full model with interactions
ordinal_model_s2 <- brm(
  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = "s2_ordinal_full6.rds"
)

@@

@@

odds_ratio_summary_s2
```

            comparison odds_ratio ci_lower ci_upper
1    Percentage vs kWh       3.31     0.43     11.3
2           USD vs kWh       6.64     0.84     25.2
3       Rounded vs Not       0.41     0.09      1.1
4 15% Goal vs 10% Goal       0.51     0.12      1.5

@@

# Fit the full model with interactions
ordinal_model_s2 <- brm(
  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = "s2_ordinal_full6.rds"
)

@@

odds_ratio_summary_s2

posterior_samples_s2 <- as.data.frame(ordinal_model_s2)

odds_ratio_summary_s2 <- data.frame(
  comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
  odds_ratio = c(
    mean(odds_ratios_s2$Percentage_vs_kWh),
    mean(odds_ratios_s2$USD_vs_kWh),
    mean(odds_ratios_s2$Rounded_vs_Not),
    mean(odds_ratios_s2$Goal15_vs_10)
  ),
  ci_lower = c(
    quantile(odds_ratios_s2$Percentage_vs_kWh, 0.025),
    quantile(odds_ratios_s2$USD_vs_kWh, 0.025),
    quantile(odds_ratios_s2$Rounded_vs_Not, 0.025),
    quantile(odds_ratios_s2$Goal15_vs_10, 0.025)
  ),
  ci_upper = c(
    quantile(odds_ratios_s2$Percentage_vs_kWh, 0.975),
    quantile(odds_ratios_s2$USD_vs_kWh, 0.975),
    quantile(odds_ratios_s2$Rounded_vs_Not, 0.975),
    quantile(odds_ratios_s2$Goal15_vs_10, 0.975)
  )
)

>>>>

---
title: "Ordinal Models"
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: ordinal_hugo.md
  gfm:
    echo: true
    output-file: ordinal.md
---

@@

pacman::p_load(dplyr,purrr,tidyr,here,tibble,stringr,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())
options(brms.backend="cmdstanr",mc.cores=4)

s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
```

@@

# Get predicted probabilities
pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, .category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")

@@

odds_ratio_summary |> kable()
# |comparison        | odds_ratio| ci_lower| ci_upper|
# |:-----------------|----------:|--------:|--------:|
# |Percentage vs kWh |        2.3|     0.62|      6.2|
# |USD vs kWh        |       14.1|     4.08|     38.0|

odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")

@@

|comparison        | odds_ratio| ci_lower| ci_upper|
|:-----------------|----------:|--------:|--------:|
|Percentage vs kWh |        2.3|     0.62|      6.2|
|USD vs kWh        |       14.1|     4.08|     38.0|

## S1  posterior predictive checks

@@

bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])

@@

# Create new_data with one id and state
new_data <- expand.grid(
  refClass = c("kWh", "Percentage", "USD"),
  id = unique(s1_agg$id)[1],      # Take first id
  state = unique(s1_agg$state)[1] # Take first state
)

f_df <- as.data.frame.table(f) |>
  rename(Probability = Freq) |>
  # Extract Y levels from the Var3 column
  mutate(Y = str_extract(Var3, "(?<=P\\(Y = ).*(?=\\))")) |>
  # Join with new_data using crossing instead of bind_cols
  crossing(new_data) |>
  # Select and arrange columns as needed
  select(refClass, id, state, Y, Probability)

@@

## Alt s1 visual of predicted probs

@@

@@

@@

## s1 probit

@@

# compare
rstanarm::loo_compare(loo_logit, loo_probit)

@@

pred_data <- data.frame(
  refClass = c("kWh", "Percentage", "USD"),
  id = unique(s1_agg$id)[1],      # Take first id
  state = unique(s1_agg$state)[1] # Take first state
)

@@

#             comparison odds_ratio ci_lower ci_upper
# 1    Percentage vs kWh       1.99     0.73     4.43
# 2           USD vs kWh       3.69     1.33     8.29
# 3       Rounded vs Not       0.70     0.57     0.84
# 4 15% Goal vs 10% Goal       0.78     0.63     0.94

odds_ratio_summary_s2 |> pander::pandoc.table(caption="Study 2: Odds ratios of accuracy")

@@

## Interaction model

```{r}
#| eval: false

# Fit the full model with interactions
ordinal_model_s2 <- brm(
  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache","s2_ordinal_full6.rds"))
)

@@

# |refClass   |rounded     |pct_goal |.category      | mean_prob| lower_ci| upper_ci|
# |:----------|:-----------|:--------|:--------------|---------:|--------:|--------:|
# |kWh        |Not Rounded |10%      |Exact match    |      0.09|     0.02|     0.25|
# |kWh        |Not Rounded |10%      |0.01-2% error  |      0.20|     0.06|     0.36|
# |kWh        |Not Rounded |10%      |2.01-15% error |      0.64|     0.38|     0.77|
# |kWh        |Not Rounded |10%      |Over 15% error |      0.07|     0.01|     0.19|
# |kWh        |Not Rounded |15%      |Exact match    |      0.19|     0.05|     0.44|
# |kWh        |Not Rounded |15%      |0.01-2% error  |      0.29|     0.14|     0.41|
# |kWh        |Not Rounded |15%      |2.01-15% error |      0.49|     0.22|     0.73|
# |kWh        |Not Rounded |15%      |Over 15% error |      0.03|     0.01|     0.09|
# |kWh        |Rounded     |10%      |Exact match    |      0.22|     0.06|     0.49|
# |kWh        |Rounded     |10%      |0.01-2% error  |      0.31|     0.16|     0.41|
# |kWh        |Rounded     |10%      |2.01-15% error |      0.45|     0.18|     0.71|
# |kWh        |Rounded     |10%      |Over 15% error |      0.02|     0.00|     0.07|
# |kWh        |Rounded     |15%      |Exact match    |      0.28|     0.07|     0.61|
# |kWh        |Rounded     |15%      |0.01-2% error  |      0.32|     0.18|     0.42|
# |kWh        |Rounded     |15%      |2.01-15% error |      0.38|     0.11|     0.68|
# |kWh        |Rounded     |15%      |Over 15% error |      0.02|     0.00|     0.06|
# |Percentage |Not Rounded |10%      |Exact match    |      0.04|     0.01|     0.13|
# |Percentage |Not Rounded |10%      |0.01-2% error  |      0.12|     0.03|     0.28|
# |Percentage |Not Rounded |10%      |2.01-15% error |      0.70|     0.53|     0.78|
# |Percentage |Not Rounded |10%      |Over 15% error |      0.14|     0.03|     0.38|
# |Percentage |Not Rounded |15%      |Exact match    |      0.10|     0.02|     0.29|
# |Percentage |Not Rounded |15%      |0.01-2% error  |      0.21|     0.06|     0.37|
# |Percentage |Not Rounded |15%      |2.01-15% error |      0.63|     0.34|     0.77|
# |Percentage |Not Rounded |15%      |Over 15% error |      0.07|     0.01|     0.20|
# |Percentage |Rounded     |10%      |Exact match    |      0.15|     0.03|     0.40|
# |Percentage |Rounded     |10%      |0.01-2% error  |      0.26|     0.09|     0.39|
# |Percentage |Rounded     |10%      |2.01-15% error |      0.55|     0.24|     0.75|
# |Percentage |Rounded     |10%      |Over 15% error |      0.04|     0.01|     0.13|
# |Percentage |Rounded     |15%      |Exact match    |      0.13|     0.03|     0.35|
# |Percentage |Rounded     |15%      |0.01-2% error  |      0.25|     0.09|     0.39|
# |Percentage |Rounded     |15%      |2.01-15% error |      0.57|     0.28|     0.76|
# |Percentage |Rounded     |15%      |Over 15% error |      0.05|     0.01|     0.14|
# |USD        |Not Rounded |10%      |Exact match    |      0.02|     0.00|     0.07|
# |USD        |Not Rounded |10%      |0.01-2% error  |      0.07|     0.01|     0.19|
# |USD        |Not Rounded |10%      |2.01-15% error |      0.67|     0.42|     0.78|
# |USD        |Not Rounded |10%      |Over 15% error |      0.24|     0.06|     0.56|
# |USD        |Not Rounded |15%      |Exact match    |      0.06|     0.01|     0.18|
# |USD        |Not Rounded |15%      |0.01-2% error  |      0.14|     0.03|     0.32|
# |USD        |Not Rounded |15%      |2.01-15% error |      0.68|     0.47|     0.78|
# |USD        |Not Rounded |15%      |Over 15% error |      0.13|     0.02|     0.37|
# |USD        |Rounded     |10%      |Exact match    |      0.05|     0.01|     0.18|
# |USD        |Rounded     |10%      |0.01-2% error  |      0.13|     0.03|     0.32|
# |USD        |Rounded     |10%      |2.01-15% error |      0.68|     0.47|     0.78|
# |USD        |Rounded     |10%      |Over 15% error |      0.13|     0.02|     0.37|
# |USD        |Rounded     |15%      |Exact match    |      0.04|     0.01|     0.14|
# |USD        |Rounded     |15%      |0.01-2% error  |      0.11|     0.02|     0.28|
# |USD        |Rounded     |15%      |2.01-15% error |      0.69|     0.50|     0.78|
# |USD        |Rounded     |15%      |Over 15% error |      0.15|     0.03|     0.41|

@@

#  odds_ratio_summary_s2
#             comparison odds_ratio ci_lower ci_upper
# 1    Percentage vs kWh       3.31     0.43     11.3
# 2           USD vs kWh       6.64     0.84     25.2
# 3       Rounded vs Not       0.41     0.09      1.1
# 4 15% Goal vs 10% Goal       0.51     0.12      1.5

odds_ratio_summary_s2 |> kable()

|comparison           | odds_ratio| ci_lower| ci_upper|
|:--------------------|----------:|--------:|--------:|
|Percentage vs kWh    |       3.31|     0.43|     11.2|
|USD vs kWh           |       6.64|     0.84|     25.2|
|Rounded vs Not       |       0.41|     0.09|      1.1|
|15% Goal vs 10% Goal |       0.51|     0.12|      1.5|

```{r}
#| eval: false

@@

# Fit the full model with interactions
ordinal_model_s2 <- brm(
  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache","s2_ordinal_full6.rds"))
)

@@

odds_ratio_summary_s2
#   Percentage_vs_kWh USD_vs_kWh Rounded_vs_Not Goal15_vs_Goal10
# 1               3.3        6.6           0.41             0.51

### 2024-11-20 misc/outlier.qmd
add manuscript, and restructure website

---
title: "Outliers and discrepancies"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  # hugo-md:
  #   include: true
  #   html-math-method: mathjax
  #   output-file: results_hugo.md
  # gfm:
  #   echo: true
  #   output-file: results_gfm.md
---

>>>>

---
title: "Outliers and discrepancies"
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  # hugo-md:
  #   include: true
  #   html-math-method: mathjax
  #   output-file: results_hugo.md
  # gfm:
  #   echo: true
  #   output-file: results_gfm.md
---
